Draft Report

### 2020-11-13 DraftReport.Rmd
Updated the draft report with the introduction

2. A Model Building Process

>>>>

### 1. An Introduction

For this report, we will be investigating the question, "How is the COVID case rate related to the population density and distribution of demographics of a state? How does the effect of population make-up on case rate compare with the effect of policy decisions made in each state?". This research question aims to measure the case rate per 100,000 residents within each state based on the make-up of the population, as well as the policy decisions that were or were not put into place in order to combat the rise of COVID cases. In this sense, we aim to measure how dependent the COVID case rate is on that of which cannot be controlled (i.e., population and demographics) vs that of which can be controlled to a certain degree (i.e., implementation of policies to attempt to combat cases). With this information, we may be able to suggest whether or not the proliferation of COVID can be controlled, and to the level at which these policies help with this control.

This question can be broken up into two distinctive areas of investigation. The first and primary area of investigation we are interested in is how the COVID case rate is affected by population metrics. For this, our key variables are COVID case rate per 100,000 residents and the population density per square mile. This will give us an initial understanding of our interest in the relationship between the spread of the virus and how densely populated a state is. Following this, we will also consider variables that address the make-up of these populations, specifically focusing on the distribution of demographics such as race/ethnicity and age. In order to account for the varying population sizes across the states, we will operationalize these variables by specifically looking at variables that contain a rate, in order to standardize across the samples. In this sense, we will be quantifying the spread of COVID by leveraging the COVID case rate per 100,000 rate, and for population demographics we will be investigating the percentage distribution of the race/ethnicity and age categories (e.g., White % of Total Population).

The secondary area of investigation will be to also include variables that measure policy implementation within states, and to compare the effect of these as opposed to the primary area of investigation that is population metrics. Specifically, we will be focusing on policies that address the mandates of wearing masks within the state. Through this, we aim to analyze whether the implementation of mask related policies will have an effect on case rate, as well as the degree of this effect relative to population metrics. To operationalize these policies and simplifying the measurements, we will only be considering whether or not these policies were implemented through indicator variables (1 = implemented, 0 = not implemented).

We preface our analysis by identifying certain assumptions and assessing the appropriateness of the data. Although there may be other considerations against the appropriateness of the data, we highlight three arguments that must be taken into account when interpreting the results of our analysis.

Firstly, it is important to note that given we are treating each state (with the addition of the District of Columbia) as a data point, our sample contains 51 data points. Although we do meet the general rule of sample size, as we begin to factor in the wide range of potential population distributions and demographics within the states, we see that we have large variance our sample. For example, the population density can have large variability depending on the population concentration among a few large areas as well as the level of uninhabited or sparsely inhabited land within a state, while the demographics of population can depend on a wide range of factors such as geographic location and employment opportunities. We highlight this to show that the sample cannot be assumed to be IID.

Secondly, we note that there are many internal and external aspects of our selected variables that have not been included within our models. Just addressing the internal information that is lost, we can see that much of the information is not captured given our method of operationalization. Among others, the loss of date-related information within our policy implementation variables strips us of any contextual knowledge regarding the length of policy implementation, and precludes us from identifying how long a specific policy was implemented. If we assume that policy implementation has some effect on case rate, this difference could play a large part in the effect. There are also other external aspects that cannot be included, such as the level of population density broken out among different population demographics.

Finally, with regard to our policy variables, for this report we have focused solely on mask mandates. Given this, we do not capture the effect of other policies that may or may not have a greater effect on the case rate (e.g., implementation of stay-at-home orders). This follows for our population-related variables, as we capture race/ethnicity and age, but not other characteristics such as proportions of residents with pre-existing conditions and economic factors. In justification, these choices were largely made due to there either being an appropriate amount of samples in each category (e.g., most states have implemented basic policies such as the closure of non-essential businesses and the implementation of stay-at-home orders, so we chose not to include such policies given thelack of samples for the states that have not), or that data on other policies or demographics was not readily available in our dataset.

### 2. A Model Building Process
